API design for machine learning software: experiences from the scikit-learn project

### 2013-08-09 paper.tex

The library has been designed to tie in with the set of numeric
and scientific packages centered around the NumPy and SciPy libraries. NumPy
\citep{vanderwalt2011} augments Python with a contiguous numeric array datatype
and fast array computing primitives,
while SciPy \citep{varoquaux2013scipy} extends it
further with common numerical operations, either by implementing
these in Python/NumPy or by wrapping existing C/C{}\verb!++!/Fortran
implementations. The \textit{scikits} were created to complement SciPy with
domain-specific toolkits. Currently, the two most popular and feature-complete
ones are by far \sklearn and
\textit{scikit-image},\footnote{\url{http://scikit-image.org}}
which does image processing.

Started in 2007, \sklearn is developed by an international team of a dozen of
core developers, mostly researchers from various fields (e.g.,
computer science, neuroscience, astrophysics). The project also benefits
from many occasional contributors proposing small bugfixes or
improvements. Development proceeds on GitHub\footnote{\url{https://github.com/scikit-learn}},
a platform which greatly facilitates this kind of
collaboration. Because of the large number of developers, emphasis is
put on keeping the project maintainable. In particular, code must follow
specific quality guidelines, such as style consistency and unit-test coverage.
Documentation and examples are required for all features,
and major changes must pass code review by at least two
developers not involved in the implementation of the proposed change.

@@

A distinguishing feature of the \sklearn API is its ability to
compose new estimators from several base estimators. Composition mechanisms can
be used to combine typical machine learning workflows into a single object which
is itself an estimator, and can be employed wherever usual estimators can be used.
In particular,
they integrate within model selection routines, allowing one to optimize at once
every step of a complex workflow. Composition of estimators can be done in two
ways: either sequentially through \texttt{Pipeline} objects or in a parallel
fashion through \texttt{FeatureUnion} objects.

@@

While part of the \sklearn API is necessarily Python-specific,
core concepts may be applicable to
machine learning applications and toolkits
written in other (dynamic) programming languages.
The power, and extensibility of the \sklearn API is evidenced
by the large and growing user-base, its use to solve real
problems across a wide array of fields,
as well as the appearance of third-party packages
that follow the \sklearn conventions. Examples of such packages include
\textit{astroML}\footnote{\url{http://astroml.org/}}
\citep{vanderplas2012astroML}, a package providing
machine learning tools for astronomers and
\textit{wiseRF}\footnote{\url{http://wise.io}}, a commercial random forest
implementation. The source code of
the recently-proposed sparse multiclass
algorithm of \citet{mblondel-mlj2013}, released as
part of the
\textit{lightning}\footnote{\url{https://github.com/mblondel/lightning}}
package, also follows the \sklearn conventions.
To maximize ease of use, we encourage more researchers
to follow these conventions when releasing their software.

>>>>

The library has been designed to tie in with the set of numeric
and scientific packages centered around the NumPy and SciPy libraries. NumPy
\citep{vanderwalt2011} augments Python with a contiguous numeric array datatype
and fast array computing primitives,
while SciPy \citep{varoquaux2013scipy} extends it
further with common numerical operations, either by implementing
these in Python/NumPy or by wrapping existing C/C{}\verb!++!/Fortran
implementations. Building upon this stack, a series of libraries called
\textit{scikits} were created, to complement SciPy with
domain-specific toolkits. Currently, the two most popular and feature-complete
ones are by far \sklearn and
\textit{scikit-image},\footnote{\url{http://scikit-image.org}}
which does image processing.

Started in 2007, \sklearn is developed by an international team of over a dozen
core developers, mostly researchers from various fields (e.g.,
computer science, neuroscience, astrophysics). The project also benefits
from many occasional contributors proposing small bugfixes or
improvements. Development proceeds on GitHub\footnote{\url{https://github.com/scikit-learn}},
a platform which greatly facilitates this kind of
collaboration. Because of the large number of developers, emphasis is
put on keeping the project maintainable. In particular, code must follow
specific quality guidelines, such as style consistency and unit-test coverage.
Documentation and examples are required for all features,
and major changes must pass code review by at least two
developers not involved in the implementation of the proposed change.

@@

A distinguishing feature of the \sklearn API is its ability to
compose new estimators from several base estimators. Composition mechanisms can
be used to combine typical machine learning workflows into a single object which
is itself an estimator, and can be employed wherever usual estimators can be used.
In particular, \sklearn's model selection routines
can be applied to composite estimators, allowing global optimization
of all parameters in a complex workflow.
Composition of estimators can be done in two
ways: either sequentially through \texttt{Pipeline} objects, or in a parallel
fashion through \texttt{FeatureUnion} objects.

@@

While part of the \sklearn API is necessarily Python-specific,
core concepts may be applicable to
machine learning applications and toolkits
written in other (dynamic) programming languages.
The power, and extensibility of the \sklearn API is evidenced
by the large and growing user-base, its use to solve real
problems across a wide array of fields,
as well as the appearance of third-party packages
that follow the \sklearn conventions. Examples of such packages include
\textit{astroML}\footnote{\url{http://astroml.org/}}
\citep{vanderplas2012astroML}, a package providing
machine learning tools for astronomers, and
\textit{wiseRF}\footnote{\url{http://wise.io}}, a commercial random forest
implementation. The source code of
the recently-proposed sparse multiclass
algorithm of \citet{mblondel-mlj2013}, released as
part of the
\textit{lightning}\footnote{\url{https://github.com/mblondel/lightning}}
package, also follows the \sklearn conventions.
To maximize ease of use, we encourage more researchers
to follow these conventions when releasing their software.
